Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-04-03 methods/subsections/machine learning.tex
Updates from Overleaf

\section{Introduction}
Machine Learning (ML), and particularly Deep Learning, has significantly advanced the MIR field by enabling automatic feature learning, improving classification and analysis tasks, and facilitating novel applications like music generation and source separation. Adopting musically motivated neural network architectures \cite{musicmotivCNN} blends together musical concepts and perceptual approaches and use this knowledge to analyze and understand music in a way that goes beyond traditional music analysis techniques.

@@

A Convolutional Neural Network (CNN), is a deep-learning neural network. The architecture of a CNN is composed of several layers, including:

@@

\item \textbf{Pooling layers:} These layers downsample the data, reducing the spatial or temporal dimensions of the input while retaining important information. A common pooling operation is the max-pooling, which can be defined as follows: given an input matrix $I$ and a pooling window of size $m \times n$, the max-pooling operation selects the maximum value within each non-overlapping $m \times n$ region of the input matrix. This operation effectively reduces the input's dimensions while retaining the most important information in each region.
\vspace*{3mm}

\item \textbf{Fully connected layers:} These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features learned in the previous layers. Mathematically, a fully connected layer can be represented as $Y = XW + b$, where $X$ is the input matrix, $W$ is the weight matrix, $b$ is the bias vector, and $Y$ is the output matrix. The non-linearity is introduced through the activation function, which is applied element-wise to the output matrix, such as $Y' = f(Y)$, where $f$ represents the activation function and $Y'$ is the transformed output matrix.
\vspace*{3mm}

>>>>

\section{Introduction}
Machine Learning (ML), particularly Deep Learning, has significantly advanced the MIR field by enabling automatic feature learning, improving classification and analysis tasks, and facilitating novel applications like music generation and source separation. Adopting musically motivated neural network architectures \cite{musicmotivCNN} blends musical concepts and perceptual approaches and use this knowledge to analyze and understand music in a way that goes beyond traditional music analysis techniques.

@@

A Convolutional Neural Network (CNN) is a deep-learning neural network. The architecture of a CNN is composed of several layers, including:

@@

\item \textbf{Pooling layers:} These layers downsample the data, reducing the spatial or temporal dimensions of the input while retaining important information. A joint pooling operation is the max-pooling, which can be defined as follows: given an input matrix $I$ and a pooling window of size $m \times n$, the max-pooling operation selects the maximum value within each non-overlapping $m \times n$ region of the input matrix. This operation effectively reduces the input's dimensions while retaining the most essential information in each region.
\vspace*{3mm}

\item \textbf{Fully connected layers:} These layers connect every neuron in one layer to every neuron in another, allowing the network to learn non-linear combinations of the features learned in the previous layers. Mathematically, a fully connected layer can be represented as $Y = XW + b$, where $X$ is the input matrix, $W$ is the weight matrix, $b$ is the bias vector, and $Y$ is the output matrix. The nonlinearity is introduced through the activation function, which is applied element-wise to the output matrix, such as $Y' = f(Y)$, where $f$ represents the activation function and $Y'$ is the transformed output matrix.
\vspace*{3mm}

### 2023-04-03 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\section{Model architecture and training strategy}
\subsection{Contrastive Self-Supervised Learning for Musical Representations (CSSLMR)}

@@

The CSSL methodology involves the following steps:

>>>>

\section{Model architecture and training strategy}
Piaget's theory of cognitive development \cite{Piagets_Theory_of_Cognitive_Development} posits that children acquire knowledge from sensory and motor experiences in their early stages, from birth to around 18 months. This period referred to as the 'sensorimotor stage', sees the emergence of early representational thought through basic actions such as sucking, grasping, looking, and listening. As children progress through different developmental stages until adolescence and adulthood, their reasoning progressively moves toward abstract ideas and the use of deductive logic. Schemas, higher-order cognitive structures hypothesized to underlie many aspects of human knowledge and skill, emerge during this process, which Piaget interprets through an equilibration mechanism that balances new information with old knowledge. This mechanism involves \textit{assimilation} and \textit{accommodation}, which entail taking in new information to fit pre-existing schemas and modifying existing schemas resulting from new information. In the context of learning, complex structures are based on simpler ones, and transfer of learning occurs when dynamic structures capable of generalization are created. \cite{audioselfsupsurvey}

There are similarities between assimilation and accommodation in Piaget's theory of cognitive development and backpropagation in machine learning, both of which involve modifying existing structures based on new information. However, it should be noted that these processes' underlying mechanisms and contexts are distinct. While assimilation and accommodation involve modifying mental structures to incorporate new information, backpropagation entails adjusting neural network weights to improve output predictions. Assimilation and accommodation occur in cognitive development, while backpropagation is specific to machine learning. Thus, while there are broad similarities between these processes, they are fundamentally different and cannot be directly compared in detail.

\subsection{Contrastive Self-Supervised Learning for Musical Representations (CSSLMR)}

@@

The CSSLMR methodology involves the following steps:
